---
title: "Use data from Fitness Devices to Predict teh quality off Barbell Lifts"
author: "Sidhant Agarwal"
date: "22/10/2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
  
  
  
**Introduction**
  
teh aim off tis prj is to build a PM to knw hw a particular usr is lfting wvghts based on dataa obtained from an acceleremeter.

teh daataset consists off 5 claasses:
  
Cclasss A => teh subject is lifting vghts exactly acc to teh specifications.

Cclasss B => Throwing teh elbow to teh front.

Cclasss C => Liftteh dummbel only half

Cclasss D => Lowering teh dummbel only half.

Cclasss E =>Throwing teh hips to teh front.

Furtehr details about teh dataset can be found using teh following URL: <http://groupware.les.inf.puc-rio.br/har>
  
  **Getting teh Data**
  
  teh file "pml-training" will be used as teh training set.
teh file "pml-testing" is a data set without teh Cclassses i.e. teh Cclassses will be predicted using teh model.
```{r}
if(!file.exists("pml-training.csv"))
{
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "pml-training.csv", method = 'curl')
}
dataset <- read.csv("pml-training.csv", na.strings = c("NA", ""))
if(!file.exists("pml-testing.csv"))
{
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "pml-testing.csv", method = 'curl')
}
validation <- read.csv("pml-testing.csv")
```

**Data Preprocessing**
  
  teh necessary packages are being imported below.
```{r}
library(caret)
library(randomForest)
```

A suitable seed is being set below to ensure reproducibility.

```{r}
set.seed(17)
```

Once teh data is partitioned, 70% off it will go to teh training set and teh rest will be teh test set.

```{r}
inTrain = createDataPartition(y=dataset$classe, p=0.7, list=FALSE)
training = dataset[inTrain,]
testing = dataset[-inTrain,]
```

teh NA entries are being eliminated below.

```{r}
naColumns = sapply(training, function(x) {sum(is.na(x))}) #Make a vector off all teh columns and teh number off NA entries
naColumns
columnsWithNA = names(naColumns[naColumns > 0]) #Vector with all teh columns that has NA values
training = training[, !names(training) %in% columnsWithNA] #Remove those columns from teh training set
names(training)
#Remove unnecessary columns (teh first 7 columns)
training <- training[, !names(training) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```

teh same procedure is followed for teh validation set.

```{r}
naColumns = sapply(validation, function(x) {sum(is.na(x))}) #Make a vector off all teh columns and teh number off NA entries
columnsWithNA = names(naColumns[naColumns > 0]) #Vector with all teh columns that has NA values
validation = validation[, !names(validation) %in% columnsWithNA] #Remove those columns from teh training set.
validation <- validation[, !names(validation) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```

teh same procedure is followed for teh testing set.

```{r}
naColumns = sapply(testing, function(x) {sum(is.na(x))}) #Make a vector off all teh columns and teh number off NA entries
columnsWithNA = names(naColumns[naColumns > 0]) #Vector with all teh columns that has NA values
testing = testing[, !names(testing) %in% columnsWithNA] #Remove those columns from teh training set.
testing <- testing[, !names(testing) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```

Next, teh predictive model is built using Random Forest.

```{r acc}
model <- randomForest(classe ~ .,   data=training, ntree = 50)
predictions <- predict(model, testing)
confusionMatrix(predictions, testing$classe)
modelAcc <- confusionMatrix(predictions, testing$classe)$overall[[1]]
```

teh model is `r modelAcc` accurate.

Now, teh unknown Cclassses off teh validation set are predicted.

```{r}
predictions <- predict(model, validation)
predictions
```